### 2023-04-03 outputs/paper/paper.qmd
Correcting fig-3

---
title: "This is my paper title"
subtitle: "This is my subtitle"
author: 
  - Jueun Kang
thanks: "Code and data are available at: LINK"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Quarto workspace setup and reading csv's into quarto ####
library(tidyverse)
library(palmerpenguins)
library(haven)
library(here)
library(dplyr)

# respondents' degree obtained 
clean_data_degree <- read_csv("~/Documents/environment_education/inputs/data/degrees_clean.csv")

# respondents' willingness to donate towards protecting the environment
clean_data_willingness <- read_csv("~/Documents/environment_education/inputs/data/willingness_clean.csv")

# respondents who donated to an environmental group
clean_data_donation <- read_csv("~/Documents/environment_education/inputs/data/donation_clean.csv")

```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# Data {#sec-data}

Our data is of penguins (@fig-bills).

## Data Source and Methodology

Talk about data source and methodology

### Strengths and Weaknesses of GSS

Talk about strengths and weaknesses of GSS

### Potential Ethical and Bias Issues

Talk about potential ethical and bias issues

### Attributes

Attributes go here

## Exploration

```{r}
#| label: fig-1
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

clean_data_degree$degree <-
  factor(degrees_clean$degree,
         levels = c("< High School", "High School", "Junior College", "Bachelors", "Graduate"))

degree_obtained |>
  ggplot(aes(x = degree, y = freq)) +
  geom_col(fill = "slateblue1") +
  theme_minimal() +
  labs(x = "Degree Obtained",
       y = "Number of Respondents",
       title = "Degree obtained by respondents")
```

@fig-1 talk more about it.

```{r}
#| label: fig-2
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

clean_data_willingness$grnprice <-
  factor(willingness_clean$grnprice,
         levels = c("Very Unwilling", "Fairly Unwilling", "Neutral", "Fairly Willing", "Very Willing"))

clean_data_willingness |>
  ggplot(aes(x = grnprice)) +
  geom_bar(fill = "slateblue1") +
  theme_minimal() +
  labs(x = "Willingness to donate",
       y = "Number of respondents",
       title = "Willingness to donate to environmental protection")
```
@fig-2 talks about.

# Results

```{r}
#| label: fig-3
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

degree_and_willingness <-
  clean_data_degree |>
  left_join(clean_data_willingness, by = join_by(id), suffix = c("Degree", "Willingness"))

degree_and_willingness |>
  ggplot(aes(x = degree, y = grnprice)) +
  geom_jitter() +
  theme_minimal() +
  labs(x = "Degree Obtained by Respondents",
       y = "Willingness to Pay Towards Environmental Protection",
       title = "Respondents' Degree and Willingness to Pay Towards Environment")

```
@fig-4 talks about.

```{r}
#| label: fig-4
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

#clean_data |>
#  ggplot(mapping = aes(x = degrees_clean, fill = grnmoney)) +
# geom_bar(position = "dodge2")

```

# Discussion

textbook 5.2.1

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References